\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-19 paper.qmd
Added plan for vis for tuning
Added intro to simulated data example

@@

@@

@@

@@

>>>>

<!--add distribution of density and add the first quartile as the benchmark-->

@@

<!--add vis showing removing the hexagons using first quantile, and observing neighbors-->

@@

<!--add MSE vs density (0 to 1)-->

@@

<!--add distribution of distance along with the default benchmark-->
<!--show the trimesh with colored long edges-->

@@

In this section, the effectiveness of the algorithm is described using a simulated dataset. The dataset consists of five spherical Gaussian clusters in $4\text{-}D$, with each cluster containing an equal number of points and the same within-cluster variation.

In the 2D layouts generated by various NLDR techniques, as shown in @fig-NLDR-variety-gau, five well-separated clusters are shown. In tSNE (see @fig-NLDR-variety-gau (a)), these clusters appear closely. UMAP arranges all clusters in a parallel manner, with three aligned in one line and the other two in a separate line (see @fig-NLDR-variety-gau (b)). In contrast, PHATE shows two closely positioned clusters and three more distant ones (see @fig-NLDR-variety-gau (c)). In TriMAP, two clusters are close, though not as tightly as PHATE, while the other three are well-separated (see @fig-NLDR-variety-gau (d)). Finally, PaCMAP shows one central cluster and the remaining four spread out in different directions (see @fig-NLDR-variety-gau (e)).

<!-- five Gaussian clusters with different NLDR techniques-->
```{r}
#| warning: false
#| echo: false

## Import data
df_2 <- read_rds("data/five_gau_clusters/data_five_gau.rds")
training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")
test_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_test.rds")

tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
umap_data_gau <- read_rds("data/five_gau_clusters/umap_data_five_gau.rds")
phate_data_gau <- read_rds("data/five_gau_clusters/phate_data_five_gau.rds")
trimap_data_gau <- read_rds("data/five_gau_clusters/trimap_data_five_gau.rds")
pacmap_data_gau <- read_rds("data/five_gau_clusters/pacmap_data_five_gau.rds")

## Visualize embedding

nldr_gau1 <- tsne_data_gau |>
  ggplot(aes(x = tSNE1,
             y = tSNE2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("a")

nldr_gau2 <- umap_data_gau |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("b")

nldr_gau3 <- phate_data_gau |>
  ggplot(aes(x = PHATE1,
             y = PHATE2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("c")

nldr_gau4 <- pacmap_data_gau |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("d")

nldr_gau5 <- trimap_data_gau |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2)) +
  geom_point(alpha=0.1, size=1) +
  interior_annotation("e")
```

```{r}
#| label: fig-NLDR-variety-gau
#| echo: false
#| fig-cap: "Five different NLDR representations of the same data. Different techniques and different parameter choices are used. Is there a best representation of the original data or are they all providing  equivalent information?"
#| fig-width: 8
#| fig-height: 2
#| out-width: 100%
#| fig-pos: H
# (a) tSNE (perplexity = 61), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2)
nldr_gau1 + nldr_gau2 + 
  nldr_gau3 + nldr_gau4 + nldr_gau5 + 
  plot_layout(ncol = 5)
```

<!--add trimesh for all five-->
<!--add three screenshots from each technique with youtube links-->

### 2024-05-19 paper.tex
Added intro to simulated data example

>>>>

In this section, the effectiveness of the algorithm is described using a
simulated dataset. The dataset consists of five spherical Gaussian
clusters in \(4\text{-}D\), with each cluster containing an equal number
of points and the same within-cluster variation.

In the 2D layouts generated by various NLDR techniques, as shown in
Figure~\ref{fig-NLDR-variety-gau}, five well-separated clusters are
shown. In tSNE (see Figure~\ref{fig-NLDR-variety-gau} (a)), these
clusters appear closely. UMAP arranges all clusters in a parallel
manner, with three aligned in one line and the other two in a separate
line (see Figure~\ref{fig-NLDR-variety-gau} (b)). In contrast, PHATE
shows two closely positioned clusters and three more distant ones (see
Figure~\ref{fig-NLDR-variety-gau} (c)). In TriMAP, two clusters are
close, though not as tightly as PHATE, while the other three are
well-separated (see Figure~\ref{fig-NLDR-variety-gau} (d)). Finally,
PaCMAP shows one central cluster and the remaining four spread out in
different directions (see Figure~\ref{fig-NLDR-variety-gau} (e)).

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-NLDR-variety-gau-1.pdf}

\caption{\label{fig-NLDR-variety-gau}Five different NLDR representations
of the same data. Different techniques and different parameter choices
are used. Is there a best representation of the original data or are
they all providing equivalent information?}
